---
title: "Saurav Rahman_final_project_ADJ25"
output: html_notebook
---

#Load library
```{r}

library(tidyverse)

```
#Load data
#Process: I imported the data with the name "wage_theft" using the "read_csv" code.
```{r}

wage_theft <- read_csv("data/whd_whisard.csv")

```

#Get to know the data/ Data dictionary:
1. How many rows?
#Answer: 357,269 rows

2. What is one row?- 
#Answer: Each row represents a single, concluded compliance action (investigation or case) conducted by the Department of Labor at a specific employer. 

3. How many column?- 
#Answer: Initially, the dataset contained 110 columns. However, several columns contained only zero values and no meaningful data, so I excluded those. After removing these columns, the number of columns was reduced to 94. Next, I extracted the 'finding_start_date' column into separate columns for Date, Month, and Year to enable more detailed time-based analysis. This transformation increased the total number of columns to 97, with the addition of the new 'Date', 'Month', and 'Year' columns. 

4. What is in every column?
#Answer: The data set begins with the case ID, followed by the employer’s name, address, and the case start and end dates. In addition to these demographic details, there are nearly 100 columns containing data related to the number of violations, back wage amounts, and fines. From these, the 20 most relevant indicators will be selected for analysis.

#Exclude the columns that contain only zeros in all rows.

#Process: I used the select() function to choose which columns to keep, combined with where(~!all(. == 0)) to filter out columns where all values are zero. The code ~!all(. == 0), means "keep columns where not all values are zero". It removes only columns where every single value is zero. The result was stored back in the original dataset called MO_wage_theft.
```{r}

cleaned_wage_theft <- wage_theft %>%
  select(where(~!all(. == 0)))

```


#Process: At this point in my analysis, I wanted to separate the month and year from the findings_start_date column to perform time-based analysis. Initially, I used the mdy() function to convert the findings_start_date into separate Date, Month, and Year components. However, I realized that the actual date format was "ymd", so I switched to using the ymd() function to properly extract the Month and Year. While the function successfully separated the data, it returned a warning message. After doing some research, I found that adding the argument quiet = TRUE would suppress the warning. Once I included this, I was able to separate the column cleanly without any further warnings.With these updates, I created a new dataset called US_wage_theft.I chose to analyze the findings_start_date because it represents the point at which the Wage and Hour Division first determines whether there is a violation or no violation in a case. This date marks the beginning of the findings phase, where initial conclusions are drawn based on available evidence and analysis.
```{r}

US_wage_theft <- cleaned_wage_theft %>%
mutate (Date = mdy (findings_start_date), Month = month (findings_start_date), Year = year (findings_start_date))

US_wage_theft <- cleaned_wage_theft %>%
mutate (Date = ymd (findings_start_date), Month = month (findings_start_date), Year = year (findings_start_date))

#Final code: 
US_wage_theft <- cleaned_wage_theft %>%
  mutate(
    Date = ymd(findings_start_date, quiet = TRUE),
    Month = month(ymd(findings_start_date, quiet = TRUE)),
    Year = year(ymd(findings_start_date, quiet = TRUE)))

```

5.What are the date boundaries of the data? 
#need attention 
```{r}

US_wage_theft %>% 
  mutate(Date = mdy (Date)) %>% 
  summarise(range(Date))

US_wage_theft %>% 
  mutate(Date = ymd (Date)) %>% 
  summarise(range(Date))

```

6. In total, how many years and months of data are in this data set?
#Answer: I assume there are some errors in the date data. According to the data source, the period starts from 1985. However, it also includes data from the 1900s and 1960s. Additionally, there are 21 cases without a date. Despite these errors, I consider them minimal and have analyzed the data accordingly. The dataset spans a total of 436 months, from January 1990 to March 2025.

#Process: To find the total number of months and years, I used the count function, which provides month-wise data for each year.
```{r}

US_wage_theft %>% 
  count(Year, Month)

```

7. Are there any blank rows in the "case_id" column?
#Answer: No, there are no blank or missing values in the "case_id" column. Each row in the dataset contains a valid case ID.

#Process: To ensure data completeness in the main column of the dataset, I checked for any missing values in the "case_id" column using the is.na() function. This function helps identify NA values (i.e., missing data). After applying the check, I confirmed that no rows have missing values in the "case_id" column.
```{r}

US_wage_theft %>%
  filter(is.na(case_id) | case_id == "")

```

8. Are there any blank rows in the "trade_nm" (Employer name) column?
#Answer: No, the employer/business name is mentioned in all rows. There are no missing values in the "trade_nm" column.

#Process:I used the same is.na() function to check whether any employer names were missing in the "trade_nm" column. The result showed 0 (zero) missing rows, which means no data is missing in this column either.
```{r}

US_wage_theft %>%
  filter(is.na(trade_nm) | trade_nm == "")

```

9. Are there any blank rows in the "date" column?
#Answer: Yes, there are 21 rows where the finding start date is not mentioned.

#Process:I used the same is.na() function to check whether any values were missing in the "date" column. The result showed 21 missing rows, indicating that the start date is not provided in those entries.
```{r}

US_wage_theft %>%
  filter(is.na(Date) | Date == "")

```

10. Are there any blank rows in the "st_cd" (State name) column?
#Answer: Yes, there are 3 rows where the state name of the employer/business is not mentioned.

#Process:I used the same is.na() function to check for missing values in the "st_cd" column. The result showed 3 blank rows, indicating that the state name is not provided in those rows.
```{r}

US_wage_theft %>%
  filter(is.na(st_cd) | st_cd == "")

```

11. How many unique cases are in the data?
#Answer: 357,269. This indicates that there are no duplicate case ID numbers in the dataset—all case IDs are unique.

#Process: I wanted to check how many unique case_id exist and whether any case IDs are repeated or duplicated. For that, I used the summarize() function with the n_distinct() code. The result returned the same number of rows as in the dataset, which confirms that all case_id values are unique.
```{r}

US_wage_theft %>%
  summarise(case_id = n_distinct(case_id))

```

11. How many unique employers are in the data?
#Answer: 298,751. This indicates that many employers appear multiple times for wage violations in different states across the U.S.

#Process:I wanted to check how many unique employers exist in the dataset. For that, I used the summarize() function with the n_distinct() code, similar to the previous question. 
```{r}

US_wage_theft %>%
  summarise(Unique_Employer = n_distinct(trade_nm))

```

Q1. Which state has the highest number of wage theft cases in the United States, according to Wage and Hour Compliance Action Data? Where does Missouri rank on the list?
#Answer: Texas has the highest number, with 45,983 individual cases of Wage and Hour Compliance violations. Missouri ranks 13th on the list, with 7,871 violation cases.

#Process: To identify the state-wise wage theft cases, I used the count() function. Since each case is represented by an individual row, and each row includes a specific state name, I counted the state using the count() function. Then, I arranged the results in descending order to find the state with the highest number of wage theft cases in the U.S. 
```{r}

US_wage_theft %>%
  count(st_cd) %>%
  arrange (desc (n))


```

Q2. Which state experienced the highest total number of wage theft violations in the United States, according to Wage and Hour Compliance Action Data? Where does Missouri rank on the list?
#Answer: Florida recorded the highest number, with a total of 1,185,054 Wage and Hour Compliance violations across 29,046 cases. Missouri ranks 25th on the list, with a total of 131,087 violations.

#Process:The Department of Labor found several compliance violations in each wage theft case, meaning that a single case can involve multiple violations. Based on these violations, authorities fined the companies. Therefore, it is important to examine the number of violations committed by businesses in different states.To do this, I first grouped the data by state, then summarized the case_violtn_cnt column, which represents the total number of violations per case. Finally, I arranged the results in descending order to identify the states where businesses violated wage and hour compliance the most.  
```{r}

US_wage_theft %>%
  group_by(st_cd) %>%
  summarise (Total_Violation = sum (case_violtn_cnt)) %>%
  arrange (desc (Total_Violation))

```

Q3. Against which company or business did employees file the highest number of wage theft cases in the United States, according to Wage and Hour Compliance Action Data?
#Answer: Subway. There are a total of 1,359 cases across all states in which Subway was found guilty of violating Wage and Hour Compliance.

#Process:During the assessment of the dataset, I found that there are several employers who have multiple cases across the U.S. Many of these employers operate chain businesses in different states. So, while analyzing the national data, I wanted to identify which employers or businesses have the highest number of total cases.To do this, I used the count() function on the employer name (trade_nm), similar to how I counted total cases by state. 
```{r}

US_wage_theft %>%
  count(trade_nm) %>%
  arrange (desc (n))

```

Q4. Which company or business committed the highest number of wage theft violations in the United States, according to Wage and Hour Compliance Action Data?
#Answer: Wells Fargo Bank, N.A. accounted for the highest number, with a total of 530,002 Wage and Hour Compliance violations in the U.S.

#Process: Since the number of cases and the number of violations are different, I wanted to identify which employers had the highest number of violations. To do this, I grouped the data by trade_nm (employer name), then summarized the case_violtn_cnt (total number of violations), and arranged the results in descending order. I could have included this analysis in the previous question as well. However, that would have required choosing just one indicator to sort by. In doing so, I wouldn't have been able to identify the employers with the highest numbers for both indicators, as the employers with the most cases are not necessarily the same as those with the most violations.
```{r}

US_wage_theft %>%
  group_by(trade_nm) %>%
  summarise (Total_Violation = sum (case_violtn_cnt)) %>%
  arrange (desc (Total_Violation))

```

Q5. Which employer/business had to pay the highest amount in back wages to its employees for violating Hour and Wage Compliance across the U.S.?
#Answer: Puerto Rico Department of Corrections — $39,950,933.10.

#Process:"Back Wages Agreed to Pay" is an important indicator, as it reflects the financial restitution provided to workers who were victims of wage theft or other labor violations. Therefore, I wanted to find out which employers had to pay the highest amount in back wages.To analyze this, I followed a similar approach as in the previous question. However, this time I summarized the bw_atp_amt (back wages agreed to pay) and arranged the results in descending order. The outcome revealed some new employer names that were not at the top in the previous analysis, indicating that different employers are associated with different types and levels of violations. Furthermore, I wanted to see how many employees were set to receive those back wages. So, I added another sum() function for the ee_atp_cnt indicator, which refers to the total number of employees receiving back wages. This result appeared in a separate column alongside the total back wages. 
```{r}

US_wage_theft %>%
  group_by(trade_nm) %>%
  summarise (total_Backwages = sum (bw_atp_amt), Total_Employee_to_Pay = sum (ee_atp_cmt)) %>%
  arrange (desc (total_Backwages))

```

Q6. Which employer/business had to pay the highest total amount of civil money penalties (CMP) for violating Hour and Wage Compliance across the U.S.?
#Answer: The results contain some errors and need further review.
```{r}

US_wage_theft %>%
  group_by(trade_nm) %>%
  summarise (total_cmp_assd = sum (cmp_assd)) %>%
  arrange (desc (total_cmp_assd))

US_wage_theft %>% filter(trade_nm == "Peri & Sons Farm, Inc.")

```

Q7. Which employer/business violated Hour and Wage Compliance against the highest number of employees in the U.S.?
#Answer: Walmart. A total of 88,278 employees were affected by Walmart's violations of Hour and Wage Compliance, which is the highest number in U.S.

#Process: Another important variable in this dataset is ee_violtd_cnt, which refers to the number of employees affected by violations. Using this variable, I wanted to identify which employers affected the highest number of employees across the U.S. To analyze this, I grouped the data by trade_nm (employer name) and summarized the "ee_violtd_cnt" variable. Then, I arranged the results in descending order to highlight the employers with the greatest number of affected employees.  
```{r}

US_wage_theft %>%
  group_by(trade_nm) %>%
  summarise (total_affected_employees = sum (ee_violtd_cnt)) %>%
  arrange (desc (total_affected_employees))

```

Q8. What is the year-wise trend for Wage & Hour Compliance violation cases in the U.S.?
#Answer: The year 2010 recorded the highest number of cases, with 24,954. However, the analysis shows that the period from 2009 to 2013 saw the highest yearly case counts for Wage & Hour Compliance violations. The total number of violations varied across these years. For example, in 2012, the total number of cases was comparatively lower among the top five years, but the total number of violations was the highest. 

#Process: I wanted to analyze the year-wise trend of total Wage & Hour Compliance violation cases and the total number of violations. To do this, I grouped the data by year and then summarized both the number of cases and the total violations.For the number of cases, I used the count() function with n() to count each case as a single, unique instance. For violations, I used sum() to calculate the total number of violations across all cases for each year. Finally, I arranged the results of total cases in descending order to observe which years had the highest numbers of cases.
```{r}

US_wage_theft %>%
  group_by(Year) %>%
  summarise ( count = n(), Total_Violation = sum (case_violtn_cnt)) %>%
  arrange (desc (count))

```

Q9. What is the trend in Wage & Hour Compliance violation cases in the U.S. over the last 10 years?
#Answer: Wage & Hour Compliance violation cases across the U.S. have been steadily decreasing over the past 10 years. In 2015, the total number of cases was 18,500, which dropped significantly to just 717 in 2024. However, the total number of violations within these cases varied by year. Despite these fluctuations, the number of violations has also shown a downward trend since 2019.

#process: For this question, I followed a similar approach as in the previous questions, but added a filter for specific years. To apply the filter, I used the %in% c() syntax for accurate results. 
```{r}

US_wage_theft %>%
  filter(Year %in% c(2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024)) %>%
  group_by(Year) %>%
  summarize(count = n(),
    total_violation = sum(case_violtn_cnt)) %>%
  arrange(desc(count))

```

Q10. Which industry has the highest number of Wage & Hour Compliance cases and violations in the U.S.? How many employees were affected in those industries?
#Answer: The restaurant and hotel industry has the highest number of Wage & Hour Compliance violation cases. This industry also tops the list in terms of total violations and the total number of affected employees.

#Process: To identify industry-wise wage theft patterns, I grouped the dataset by naics_code_description, which represents the name of the industry. After grouping, I summarized three key indicators: total cases, total number of violations, and total affected employees. I used count with n() to calculate the total number of wage theft cases per industry, treating each row as a unique instance. For the other two variables, I used the summarize() function to calculate the totals. Finally, I arranged the results in descending order based on the number of cases (count) to identify the industries with the highest number of reported wage and hour violations.
```{r}

US_wage_theft %>%
  group_by( industry = naics_code_description) %>%
  summarise ( count = n(), total_violation = sum (case_violtn_cnt), total_affected_employees = sum (ee_violtd_cnt)) %>%
  arrange (desc (count))

```

Q11: Which indicators of Wage & Hour Compliance Action data has the highest total violation? 
#Answer: Family and Medical Leave Act (FMLA) violation : 6,618,226. folwing Service Contract Violation: 480,704. 

#Process: Wage & Hour Complince  Action data has more than 80 variable, half of those are relate to voilation count. Among them 6 veriable are most impotant to follow accoding to Department of Labor. That why I messure those to answer this question. First, I selected six specific violation-related columns from the dataset: flsa_violtn_cnt (Fair Labor Standards Act violations), mspa_violtn_cnt (Migrant and Seasonal Agricultural Worker Protection Act violations), sca_violtn_cnt (Service Contract Act violations), fmla_violtn_cnt (Family and Medical Leave Act violations), h1b_violtn_cnt (H-1B work visa violations), flsa_cl_violtn_cnt (Child Labor violations under FLSA). Then, I used the summarize() function to calculate the total number of violations under each law. I renamed the output columns for clarity, so each total reflects the respective violation type. However, the way the results were displayed using the previous function made it difficult to measure or analyze the data clearly. Therefore, I wanted to restructure the output by having all violation types listed in one column and their corresponding violation numbers in another column. To achieve this, I used the pivot_longer() function. In my mid-term project, I had used pivot_wider(), so I was already familiar with how the pivot_longer() function works. I applied it here to display the results in a more readable format and then arranged them in descending order for easier comparison and analysis.
```{r}

US_wage_theft %>%
    select(flsa_violtn_cnt, mspa_violtn_cnt,  sca_violtn_cnt, 
         fmla_violtn_cnt, h1b_violtn_cnt, flsa_cl_violtn_cnt) %>%
  summarize (Fair_Labor_Standards_Violation = sum (flsa_cl_violtn_cnt), 
             Migrant_Agricultural_Protection_Violation = sum (mspa_violtn_cnt), 
             Service_Contract_Violation = sum (sca_violtn_cnt),
             Family_Medical_Leave_Violation = sum (fmla_violtn_cnt), 
             Immigration_Work_Visa_Violation = sum (h1b_violtn_cnt), 
             Child_labor_Violation = sum (flsa_cl_violtn_cnt))

#Final code: 
US_wage_theft %>%
    select(flsa_violtn_cnt, mspa_violtn_cnt,  sca_violtn_cnt, 
         fmla_violtn_cnt, h1b_violtn_cnt, flsa_cl_violtn_cnt) %>%
  summarize (Fair_Labor_Standards_Violation = sum (flsa_cl_violtn_cnt), 
             Migrant_Agricultural_Protection_Violation = sum (mspa_violtn_cnt), 
             Service_Contract_Violation = sum (sca_violtn_cnt),
             Family_Medical_Leave_Violation = sum (fmla_violtn_cnt), 
             Immigration_Work_Visa_Violation = sum (h1b_violtn_cnt), 
             Child_labor_Violation = sum (flsa_cl_violtn_cnt)) %>%
  pivot_longer(cols = everything(), names_to = "Violation_Type", values_to = "Total_Count") %>%
  arrange(desc(Total_Count))

```

Q12: How has the violation number of the major indicators of Wage & Hour Compliance Action data varied over the years?
#Answer:The results show that every major indicator has decreased significantly over the past ten years. For instance: Fair Labor Standards Violation dropped from 3,550 in 2015 to just 42 in 2024. Service Contract Violation was 22,289 in 2015 and declined to 200 in 2024. Family and Medical Leave Violation decreased from 220,774 to 2,872. Child Labor Violation also saw a reduction, from 3,550 in 2015 to 42 in 2024. This trend indicates substantial improvement in wage and hour compliance over the last decade.  

#Process: To analyze year-wise data for each indicator, I first grouped the dataset by year. Since the dataset spans a long period, I filtered it to only include the most recent 10 years using the %in% operator. Then, I used the same code structure from the previous question to summarize violations for six major indicators. I also intended to calculate the negative growth percentage over the ten-year span to show the rate of improvement but was unable to write a code it successfully. Instead, I sorted the summarized data by year to observe the trend.
```{r}

US_wage_theft %>%
  filter(Year %in% c(2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024)) %>%
  group_by(Year) %>%
  select(flsa_violtn_cnt, mspa_violtn_cnt,  sca_violtn_cnt, 
         fmla_violtn_cnt, h1b_violtn_cnt, flsa_cl_violtn_cnt) %>%
  summarize (Fair_Labor_Standards_Violation = sum (flsa_cl_violtn_cnt), 
             Migrant_Agricultural_Protection_Violation = sum (mspa_violtn_cnt), 
             Service_Contract_Violation = sum (sca_violtn_cnt),
             Family_Medical_Leave_Violation = sum (fmla_violtn_cnt), 
             Immigration_Work_Violation = sum (h1b_violtn_cnt), 
             Child_labor_Violation = sum (flsa_cl_violtn_cnt)) %>%
  arrange (Year)

```
Q13: How has the violation number of the major indicators of Wage & Hour Compliance Action data varied across the states?
#Answer: The data does not provide a single, specific answer but rather a table with 57 rows representing different states and territories. However, states with the highest total number of violations also tend to rank high across various individual indicators. 

#Process: To analyze state-wise data for each indicator, I first grouped the dataset by state.Then, I used the same code structure from the previous question to summarize violations for six major indicators. 
```{r}

US_wage_theft %>%
  group_by(st_cd) %>%
    select(flsa_violtn_cnt, mspa_violtn_cnt,  sca_violtn_cnt, 
         fmla_violtn_cnt, h1b_violtn_cnt, flsa_cl_violtn_cnt) %>%
  summarize (Fair_Labor_Standards_Violation = sum (flsa_cl_violtn_cnt), 
             Migrant_Agricultural_Protection_Violation = sum (mspa_violtn_cnt), 
             Service_Contract_Violation = sum (sca_violtn_cnt),
             Family_Medical_Leave_Violation = sum (fmla_violtn_cnt), 
             Immigration_Work_Violation = sum (h1b_violtn_cnt), 
             Child_labor_Violation = sum (flsa_cl_violtn_cnt)) 

```

Q14. How the top five state that done highest violation in total, varied in terms of total violation of Wage and Hour Compliance over the last 10 years?
#Answer: All five states have seen decreases in violations by 2024. Every  state's total vilation was below thousand last year. Florida had the highest violations in 2017 with 117,685 and remained one of the top violators. Tennessee saw two spikes in violation: 266,410 in 2019 and 240,982 in 2021. Texas had high violation in early years, especially in 2015 (83,727) and 2019 (163,653), then declined. California showed consistent mid-to-high violations with a peak in 2018 (89,522). Arkansas (AR) had a high violation in 2016 (456,237) but saw a decline afterward.

#Process: To answer this question, I filtered the data by both State and Year, as I needed information for five specific states over the last ten years. From the beginning of my code, I used the %in% operator to select only the relevant states and years.Next, I grouped the data by both Year and State, since I needed the violation numbers for each state across each year. After grouping, I summarized the total number of violations for each combination of year and state. Finally, to present the data clearly in a tabular format, I used the pivot_wider() function to reshape the data—making each state a separate column with years as rows. 
```{r}

US_wage_theft %>%
  filter(st_cd %in% c("FL", "TX", "TN", "AR", "CA")) %>%
  filter(Year %in% c(2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024)) %>%
  group_by(Year, st_cd) %>%
  summarize(total_violation = sum(case_violtn_cnt)) %>%
  pivot_wider(names_from = "st_cd", 
              values_from = "total_violation") 

```

Q15. How the states varied in terms of total violation of Wage and Hour Compliance over the last 10 years?
#Answer: The data does not provide a single, specific answer but rather a table with 10 columns and 57 rows representing different states and territories. However, states with the highest total number of violations also tend to rank high across each years. However the trend of numbers for each states has been declining. 

#Process: To answer this question, I filtered the data by Year, as I needed information for ten specific years. I used the %in% operator to select only the relevant years. Next, I grouped the data by both Year and State, since I needed the violation numbers for each state across filtered year. After grouping, I summarized the total number of violations for each combination of year and state. Finally, to present the data clearly in a tabular format, I used the pivot_wider() function to reshape the data—making each state a separate column with years as rows. 
```{r}

US_wage_theft %>%
  filter(Year %in% c(2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024)) %>%
  group_by(Year, st_cd) %>%
  summarize(total_violation = sum(case_violtn_cnt)) %>%
  pivot_wider(names_from = "st_cd", 
              values_from = "total_violation") 

```
Q16. Which state take highest time to complete a findings in average? 
#Answer: Need attention

```{r}


US_wage_theft <- US_wage_theft %>%
  mutate (findings_start_date = ymd (findings_start_date), 
          findings_end_date = ymd (findings_end_date),
          Days_to_complete = as.numeric (findings_end_date - findings_start_date)) %>%
  arrange (desc(Days_to_complete))


```
#Missouri specific dataset: 

#Process: After analyzing national percspective from the Wage and Hour Compliance Action Data, I wanted to look into missouri specific data set. Therefore I filter Missouri data only for my further analysis. For that, I use filter funtion and create a new data set called "MO_wage_theft" for furthur analysis. 

```{r}
MO_wage_theft <- US_wage_theft %>%
  filter(st_cd == "MO")
```


#Missouri clean dataset
#Exclude the columns that contain only zeros in all rows.

#Process: I fllowed this process in national data set as well to exclude the rows those have any data. I used the select() function to choose which columns to keep, combined with where(~!all(. == 0)) to filter out columns where all values are zero. The code ~!all(. == 0), means "keep columns where not all values are zero". It removes only columns where every single value is zero. The result was stored back in the original dataset called MO_wage_theft with 75 columns, which was 98 for national data set.  
```{r}

MO_wage_theft <- MO_wage_theft %>%
  select(where(~!all(. == 0)))

```


#Standardize the City Names to All Uppercase Letters
#Process: I found that some city names appeared in different formats—for instance, both "Saint Louis" and "SAINT LOUIS" were present. To ensure consistency in my analysis, I standardized all city names to uppercase. To do this, I used the mutate() function along with the toupper() function. The toupper() function converts all text in the cty_nm column to uppercase letters. I updated the original dataset by assigning the result back to the same name: MO_wage_theft.
```{r}

MO_wage_theft <- MO_wage_theft %>%
  mutate(cty_nm = toupper(cty_nm))

```


Q17: Which company/employer in Missouri had the highest record of Wage & Hour Compliance cases and number of violations?
#Answer:Highest number of cases: Subway, with 33 cases. Highest number of violations: George's Processing, Incorporated, with a total of 3,148 violations across 2 cases.

#Process: To answer these two related questions, I used a similar approach in both cases, with only a small difference in the arrangement step. I wanted to present the number of cases and total violations at the same time, but since the employers with the highest case count and highest total violations were different, I wrote two separate code blocks. The codewas  simple and similar to what I used for the national-level data. First, I grouped the data by trade_nm (employer name), then summarized it. For the number of cases, I used the n() function to count how many times each employer appeared in the trade_nm column—each appearance represents a case.For the total number of violations, I used the sum() function to calculate the total violation count per employer.    
```{r}

MO_wage_theft %>%
  group_by(trade_nm) %>%
  summarise ( count = n(), 
              Total_Violation = sum (case_violtn_cnt)) %>%
  arrange (desc (count))

MO_wage_theft %>%
  group_by(trade_nm) %>%
  summarise ( count = n(), 
              Total_Violation = sum (case_violtn_cnt)) %>%
  arrange (desc (Total_Violation))


```

Q18: Which company/employer in Missouri had to pay the highest back wages, and how many employees received those back wages from each employer?
#Answer: George's Processing, Incorporated paid a total of $1,245,257.09 to 3,084 employees. 

#Process: To find the answer, I grouped the data by employer name (trade_nm). Then, I summarized the data using both bw_atp_amt (which represents back wages paid) and ee_atp_cmt (which represents the number of employees who received those wages). I arranged the summarized data in descending order by total back wages. Since the employer who paid the highest amount also happened to pay the highest number of employees, I didn’t need to write a separate code for employee count—both results could be shown in one table.
```{r}

MO_wage_theft %>%
  group_by(trade_nm) %>%
  summarise (total_Backwages = sum (bw_atp_amt), 
             Total_Employee_to_Pay = sum (ee_atp_cmt)) %>%
  arrange (desc (total_Backwages))

```

Q19: Which employer/business had to pay the highest total amount of Civil Money Penalties (CMP) in Missouri for violating Wage and Hour Compliance?
#Answer:Missouri Vegetable Farm paid the highest total amount in civil money penalties, with $384,626.61.

#Process: To find this, I grouped the data by employer name (trade_nm). After that, I summarized the cmp_assd column, which represents the total amount of civil money penalties imposed on employers. Finally, I arranged the results in descending order based on the total CMP amount (cmp_assd) to identify the top violator.  
```{r}

MO_wage_theft %>%
  group_by(trade_nm) %>%
  summarise (total_cmp_assd = sum (cmp_assd)) %>%
  arrange (desc (total_cmp_assd))

```

Q20. Which employer/business violated Hour and Wage Compliance against the highest number of employees in the Missouri?
#Answer: George's Processing, Incorporated. A total of 3147 employees were affected by  George's Processing's violations of Hour and Wage Compliance, which is the highest number in the Missouri.

#Process: To analyze this, I grouped the data by trade_nm (employer name) and summarized the "ee_violtd_cnt" variable that refers, Number of employees affected by violations. Then, I arranged the results in descending order to highlight the employers with the highest number of affected employees.  
```{r}

MO_wage_theft %>%
  group_by (trade_nm) %>%
  summarize (Total_affected_employees = sum (ee_violtd_cnt)) %>%
  arrange (desc (Total_affected_employees))

MO_wage_theft %>%
  group_by (trade_nm) %>%
  summarize (Total_affected_employees = sum (ee_violtd_cnt)) %>%
  arrange (trade_nm)

```

Q20: Which city in Missouri experienced the highest number of cases, highest number of violations, and the most affected employees according to the Wage & Hour Compliance violation data?

#Answer: Saint Louis recorded the highest numbers with 1,487 cases, 25,019 total violations, and 21,956 affected employees.

#Process: First, I grouped the data by city name (cty_nm). To answer all three parts of the question together, I used the summarize() function: For the number of cases, I used the n() function. Since each row represents a case, counting how many times a city appears gives the total number of cases. For total violations and affected employees, I used the sum() function on the relevant columns. Finally, I arranged the results in descending order by the number of cases. 
```{r}

MO_wage_theft %>%
  group_by(cty_nm) %>%
  summarise ( count = n(), 
              Total_Violation = sum (case_violtn_cnt), 
              Total_affected_employees = sum (ee_violtd_cnt)) %>%
  arrange (desc (count))

```

Q21: In which year did Missouri experience the highest number of cases, highest number of violations, and the most affected employees according to the Wage & Hour Compliance violation data?
#Answer:The highest number of cases was filed in 2009, with 527 cases. The highest number of violations occurred in 2006, with 12,329 violations. That same year (2006), 11,907 employees were affected—the highest recorded. 

#Process:The process was similar to the one I used in previous questions. I started by grouping the data by year, then used count() and summarize() to get values for the three variables: number of cases, total violations, and number of affected employees.Since the year with the most cases was different from the year with the most violations and affected employees, I used the same code but arranged the results separately—first by case count, then by total violations—to identify each highest year.
```{r}

MO_wage_theft %>%
  group_by(Year) %>%
  summarise ( count = n(), 
              Total_Violation = sum (case_violtn_cnt), 
              Total_affected_employees = sum (ee_violtd_cnt)) %>%
  arrange (desc (count))

MO_wage_theft %>%
  group_by(Year) %>%
  summarise ( count = n(), 
              Total_Violation = sum (case_violtn_cnt), 
              Total_affected_employees = sum (ee_violtd_cnt)) %>%
  arrange (desc (Total_Violation))

```

Q22: How has Missouri been experiencing Wage & Hour Compliance cases, violations, and affected employees over the last 10 years?
#Answer: Missouri has been experiencing a declining trend in Wage and Hour Compliance violations since 2020. In that year, the total number of cases was 318, which dropped significantly to  6 cases by 2024. Five years ago, the total number of violations was 3,661, and the number of affected employees was 3,066. In 2024, these numbers declined to 88 violations and 60 affected employees, respectively 

#Process: To analyze the last 10 years, I filtered the dataset using the %in% operator and listed each of the last ten years explicitly. After filtering, I followed the same approach as in the previous question: I grouped the data by year and then used count() and summarize() to calculate the number of cases, total violations, and affected employees. Finally, I arranged the data in ascending order to get year-wise trends clearly.
```{r}

MO_wage_theft %>%
  filter(Year %in% c(2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024)) %>%
  group_by(Year) %>%
  summarize(count = n(),
    total_violation = sum(case_violtn_cnt), 
    Total_affected_employees = sum (ee_violtd_cnt)) %>%
  arrange (Year)

```

Q23. Which industry in Missouri experienced the highest number of cases, highest number of violations, and the most affected employees according to the Wage & Hour Compliance violation data?

#Answer: Full-Service Restaurants recorded the highest numbers with 915 cases, 15,664 total violations, and 12,979 affected employees. Infact top three industry of the table is realted to lauiser and hosipoyality business: Full-Service Restaurants, Limited-Service Restaurants and Hotels (except Casino Hotels) and Motels.   

#Process: First, I grouped the data by industry (naics_code_description). To answer all three parts of the question together, I used the summarize() function: For the number of cases, I used the n() function. Since each row represents a case, counting how many times a industry name appears gives the total number of cases. For total violations and affected employees, I used the sum() function on the relevant columns. Finally, I arranged the results in descending order by the number of cases. 
```{r}

MO_wage_theft %>%
  group_by( industry = naics_code_description) %>%
  summarise ( count = n(), 
              total_violation = sum (case_violtn_cnt), 
              total_affected_employees = sum (ee_violtd_cnt)) %>%
  arrange (desc (count))

```

Q24. Which indicators of Wage & Hour Compliance Action data has the highest total violation in Missouri? 
#Answer: Service_Contract_Violation: 6,208. folwing Family_Medical_Leave_Violation: 2,455. 

#Process: Missouri Wage & Hour Complince data set has 75 variable, half of those are relate to voilation count. Among them 6 veriable are most impotant to follow accoding to Department of Labor. I messured those for national data as well. Those six indicators/veriables are:  flsa_violtn_cnt (Fair Labor Standards Act violations), mspa_violtn_cnt (Migrant and Seasonal Agricultural Worker Protection Act violations), sca_violtn_cnt (Service Contract Act violations), fmla_violtn_cnt (Family and Medical Leave Act violations), h1b_violtn_cnt (H-1B work visa violations), flsa_cl_violtn_cnt (Child Labor violations under FLSA). I select those for to show them in my analysis only. Then, I used the summarize() function to calculate the total number of violations under each law. I renamed the output columns for clarity, so each total reflects the respective violation type. However, the way the results were displayed using the previous function made it difficult to measure or analyze the data clearly. Therefore, I wanted to restructure the output by having all violation types listed in one column and their corresponding violation numbers in another column. To achieve this, I used the pivot_longer() function. Then arranged them in descending order for easier comparison and analysis.

```{r}

MO_wage_theft %>%
    select(flsa_violtn_cnt, mspa_violtn_cnt,  sca_violtn_cnt, 
         fmla_violtn_cnt, h1b_violtn_cnt, flsa_cl_violtn_cnt) %>%
  summarize (Fair_Labor_Standards_Violation = sum (flsa_cl_violtn_cnt), 
             Migrant_Agricultural_Protection_Violation = sum (mspa_violtn_cnt), 
             Service_Contract_Violation = sum (sca_violtn_cnt),
             Family_Medical_Leave_Violation = sum (fmla_violtn_cnt), 
             Immigration_Work_Visa_Violation = sum (h1b_violtn_cnt), 
             Child_labor_Violation = sum (flsa_cl_violtn_cnt)) %>%
  pivot_longer(cols = everything(), names_to = "Violation_Type", values_to = "Total_Count") %>%
  arrange(desc(Total_Count))

```

Q25: How has the violation number of the 6 major indicators of Wage & Hour Compliance Action data varied over the years in Missouri?
#Answer: The results show that all six major indicators have decreased significantly over the past ten years in Missouri. However, some of them experienced increases during the mid-years. For example, during the pandemic in 2020 and 2021, Fair Labor Standards Violations rose to 140, but dropped to 3 by 2024. Service Contract Violations fell from 375 in 2017 to zero in 2024. Child Labor Violations also declined, from 60 in 2015 to  3 in 2024. 
  
#Process: To analyze year-wise data for each indicator, I first grouped the dataset by year. Since the dataset spans a long period, I filtered it to only include the most recent 10 years using the %in% operator. Then, I used the same code structure from the previous question to summarize violations for six major indicators. I also intended to calculate the negative growth percentage over the ten-year span to show the rate of improvement but was unable to write a code it successfully. Instead, I sorted the summarized data by year to observe the trend.
```{r}

MO_wage_theft %>%
  filter(Year %in% c(2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024)) %>%
  group_by(Year) %>%
  select(flsa_violtn_cnt, mspa_violtn_cnt,  sca_violtn_cnt, 
         fmla_violtn_cnt, h1b_violtn_cnt, flsa_cl_violtn_cnt) %>%
  summarize (Fair_Labor_Standards_Violation = sum (flsa_cl_violtn_cnt), 
             Migrant_Agricultural_Protection_Violation = sum (mspa_violtn_cnt), 
             Service_Contract_Violation = sum (sca_violtn_cnt),
             Family_Medical_Leave_Violation = sum (fmla_violtn_cnt), 
             Immigration_Work_Violation = sum (h1b_violtn_cnt), 
             Child_labor_Violation = sum (flsa_cl_violtn_cnt)) %>%
  arrange (Year)

```

Q26. How did the top five cities in Missouri with the highest total violations of Wage and Hour Compliance over the last 10 years vary?
#Answer: Among the top five cities in Missouri that recorded the highest total violations, SPRINGFIELD, COLUMBIA, INDEPENDENCE did not have any violations of Wage and Hour Compliance in the 2024. All five cities have seen decreases in total violations by 2024 compared to 2015. However, Saint Louis recorded a higher number of violations in 2022 (312) than in the previous year (162). Columbia declined from 556 violations in 2015 to zero for the last two years. Kansas City experienced 2 violations in 2024, down from 957 in 2017. Saint Louis had the highest violations in 2017, with 1,324, and remained the top city. Independence also saw spikes in violations, reaching 1,611 in 2017. Springfield showed consistent mid-to-high violations, with a peak in 2020 (641).

#Process: To answer this question, I filtered the data by both city and year, as I needed information for five specific cities over the last ten years. From the beginning of my code, I used the %in% operator to select only the relevant years and cities. Next, I grouped the data by both year and city, as I needed the violation numbers for each city across each year. After grouping, I summarized the total number of violations for each combination of year and city. Finally, to present the data in a tabular format, I used the pivot_wider() function to reshape the data—making each city a separate column with years as rows.
```{r}

MO_wage_theft %>%
  filter(Year %in% c(2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024)) %>%
  filter(cty_nm %in% c("SAINT LOUIS", "KANSAS CITY", "SPRINGFIELD", "COLUMBIA", "INDEPENDENCE")) %>%
  group_by(Year, cty_nm) %>%
  summarize(total_violation = sum(case_violtn_cnt)) %>%
  pivot_wider(names_from = "cty_nm", 
              values_from = "total_violation") 

```

Q27. Which Wage and Hour Compliance violation case in Missouri took the longest time to complete?
#Answer: The case number is 1035967. It is against Blue Springs Lawn & Garden from Blue Springs, Missouri. This case took 4,840 days to complete, which is more than 13 years. The findings began on September 26, 1997, and ended on December 27, 2010.

#Process: Since I already created a variable called Days_to_complete that refers to the total number of days it took to close each case, I simply arranged that variable in descending order to find the answer. 
```{r}

MO_wage_theft %>%
  arrange (desc(Days_to_complete))

```


Q28. What are the average and median days to complete a Wage and Hour Compliance violation case in Missouri?
#Answer: The average number of days to complete a Wage & Hour Compliance violation case in Missouri is 589 days, and the median is 728 days, which is almost two years.

#Process: To calculate the average and median values for the days to complete, I used the summarize function and wrote code to compute both the mean (average) and median. I also added na.rm = TRUE to avoid the NA values in the column. Since the column contained a few NA values, using na.rm = TRUE ensured that the calculation was based only on the available data; otherwise, the result would have been NA.
```{r}

MO_wage_theft%>%
  summarize(average_days = mean(Days_to_complete, na.rm = TRUE),
    median_days = median(Days_to_complete, na.rm = TRUE))

```


Q29. In which industry's Wage & Hour Compliance violation cases do the median days to complete a case take the longest?
#Answer: The industry with the longest median time to complete a Wage & Hour Compliance violation case is Technical and Trade Schools. It takes 3,521 days—or approximately nine and a half years—to close a case in this industry.

#Process: To find the median days to complete a case by industry, I grouped the dataset by the naics_code_description column (which represents the industry). Then, I used the summarize() function to calculate the median of the Days_to_complete variable. Finally, I arranged the results in descending order based on the median days.  
```{r}

MO_wage_theft %>%
  group_by(naics_code_description) %>%
  summarize(median_days = median(Days_to_complete)) %>%
  arrange(desc(median_days))

MO_wage_theft %>%
  filter(Year %in% c(2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024)) %>%
  group_by(Year) %>%
  summarize(median_days = median(Days_to_complete)) %>%
  arrange (Year)

```


Q30. Which employer in Missouri violated the service contract the most, and which industry tops the list for service contract violations?
#Answer: The employer that violated the service contract the most in Missouri is Abbott Ambulance, with a total of 1,083 violations. Industry-wise, Tax Preparation Services had the highest number of service contract violations, totaling 920.

#Process: This question has two parts, so I wrote two separate pieces of code. However, both followed a similar structure, with only the grouping variable being different. To find the employer and the industry with the highest service contract violations, I used the group_by() function to group the data accordingly. Then, I used the summarize() function to calculate the total service contract violations. Finally, I arranged the results in descending order based on the number of service contract violations.
```{r}

MO_wage_theft %>%
  group_by(trade_nm) %>%
  summarize (Service_Contract_Violation = sum (sca_violtn_cnt)) %>%
  arrange (desc(Service_Contract_Violation))

MO_wage_theft %>%
  group_by(naics_code_description) %>%
  summarize (Service_Contract_Violation = sum (sca_violtn_cnt)) %>%
  arrange (desc(Service_Contract_Violation))

```

Q31. Which employer in Missouri violated the Fair Labor Standards the most, and which industry tops the list for Fair Labor Standards violations?
#Answer: The employer that violated the Fair Labor Standards the most in Missouri is Dog Days Bar and Grill, with a total of 41 violations. Industry-wise, Full-Service Restaurants had the highest number of Fair Labor Standards violations, totaling 375.

#Process: This question has two parts, similar to the previous one. The only difference is that the main variable this time is Fair Labor Standards. I summarized the data in both pieces of code by grouping first by employer and then by industry name. Finally, I arranged the results by the Fair Labor Standards violation count in decending order.

```{r}

MO_wage_theft %>%
  group_by(trade_nm) %>%
  summarize (Fair_Labor_Standards_Violation = sum (flsa_cl_violtn_cnt)) %>%
  arrange (desc(Fair_Labor_Standards_Violation))

MO_wage_theft %>%
  group_by(naics_code_description) %>%
  summarize (Fair_Labor_Standards_Violation = sum (flsa_cl_violtn_cnt)) %>%
  arrange (desc(Fair_Labor_Standards_Violation))

```

Q33. Which employer in Missouri violated the Family and Medical Leave Act the most, and which industry tops the list for Family and Medical Leave Act violations?
#Answer: The employer that violated the Family and Medical Leave Act the most in Missouri is West County Care Center, with a total of 1,380 violations. Industry-wise, Nursing Care Facilities had the highest number of Family and Medical Leave Act violations, totaling 1,421.

#Process: This question is similar to the previous two. Therefore, I followed almost the same process to answer this question, with the only change being the variable related to the Family and Medical Leave Act.
 
```{r}

MO_wage_theft %>%
  group_by(trade_nm) %>%
  summarize (Family_Medical_Leave_Violation = sum (fmla_violtn_cnt)) %>%
  arrange (desc(Family_Medical_Leave_Violation))

MO_wage_theft %>%
  group_by(naics_code_description) %>%
  summarize (Family_Medical_Leave_Violation = sum (fmla_violtn_cnt)) %>%
  arrange (desc(Family_Medical_Leave_Violation))

```


Q34. Which employer in Missouri violated the Child Labor Act the most, and which industry and city top the list for Child Labor Act violations?
#Answer: The employer that violated the Child Labor Act the most in Missouri is Dog Days Bar and Grill, with a total of 41 violations. Industry-wise, Full-Service Restaurants had the highest number of Child Labor Act violations, totaling 375. The city with the highest number of Child Labor Act violations is SPRINGFIELD, with 133 incidents.

#Process: This question is similar to the previous three, but this time I also included city as an additional factor to identify which city in Missouri experienced the most Child Labor violations. To answer all three parts of the question, I followed the same general process—grouping the data by employer, industry, and city—then summarizing the total number of Child Labor violations. The only change was using the variable related to Child_labor_Violation.
```{r}

MO_wage_theft %>%
  group_by(trade_nm) %>%
  summarize (Child_labor_Violation = sum (flsa_cl_violtn_cnt)) %>%
  arrange (desc(Child_labor_Violation))

MO_wage_theft %>%
  group_by(naics_code_description) %>%
  summarize (Child_labor_Violation = sum (flsa_cl_violtn_cnt)) %>%
  arrange (desc(Child_labor_Violation))

MO_wage_theft %>%
  group_by(cty_nm) %>%
  summarize (Child_labor_Violation = sum (flsa_cl_violtn_cnt)) %>%
  arrange (desc(Child_labor_Violation))

```

Q35. Which employer/business in Missouri violated all six major indicators of Wage and Hours Compliance?
#Answer: None.

#Process: To get the answer, I used the filter function. As I needed to find trade names of businesses that had violations in all six categories, I filtered for rows that had values greater than zero in all six columns (since zero means no violation). I wanted to see if any company had violations in all six categories, which is why I filtered using >0 for each variable. Finally, I selected the trade_nm column so that if any employer had violations in all six categories, their trade name would appear.   
```{r}

MO_wage_theft %>%
  filter(flsa_violtn_cnt > 0,
    mspa_violtn_cnt > 0,
    sca_violtn_cnt > 0,
    fmla_violtn_cnt > 0,
    h1b_violtn_cnt > 0,
    flsa_cl_violtn_cnt > 0) %>%
  select(trade_nm)

```

